State of the Practice for Medical Imaging Software Based on Open Source Repositories

### 2025-01-24 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/ResubmissionResponseToReviewer.tex

\documentclass[12pt]{casletter}

\oddsidemargin -10mm
\evensidemargin -10mm
\textwidth 180mm
\textheight 200mm
%\renewcommand\baselinestretch{0.85}

\usepackage{xcolor}

\pagestyle {plain}
\pagenumbering {arabic}

\email{smiths@mcmaster.ca}
\telephone{(905) 525-9140 ext.\ 27929}
\www{{http://}www.cas.mcmaster.ca/\~\@smiths}

\signature{Spencer Smith}

\address{Dr.\ Spencer Smith, Associate Professor\\
Computing and Software Department\\
Faculty of Engineering\\
McMaster University\\
1280 Main Street West\\
Hamilton, Ontario, Canada L8S 4K1}%  \\ Phone: 525-9140 Ext.~27929 \\E-mail:
                                  %  smiths@mcmaster.ca}

\begin{document}

\begin{letter}{~\\
    ~\\
    ~\\
    ~\\
    ~\\
    Dr.\ Anthony McGoron\\
    Editor-in-Chief\\
    Critical Reviews in Biomedical Engineering\\
    ~\\
    {\bf Re: Revisions to CRB-57015, State of the Practice for Medical Imaging
    Software Based on Open Source Repositories}}

  \opening {Dear Dr.\ McGoron:}

  Thank you and the reviewers for the feedback on our submission.  The reviewers
  provided thoughtful and constructive comments.  In response to your e-mail,
  dated January 3, 2025, we have revised the paper to incorporate the requested
  revisions.  We provide a summary below.  The revised submission also includes
  a ``diff'' version of the paper showing the additions and deletions.

  Some of the reviewer's questions are best answered by citing our related work.
  Therefore, the new version undoes the redaction originally added for the
  double blind review.

  \textbf{Reviewer \#1}

  \begin{enumerate}
  \item The relative priorities of each software quality were set equal.
  Couldn't some different weighting have been considered? Perhaps one that comes
  from interviews with developers or cited in papers? \medskip

  \emph{Papers don't agree and papers are from many domains of application for
  research software.  \cite{Nguyen-HoanEtAl2010}  Not aware of a study focused
  on medical imaging practitioners.  Interviewing a statistically significant
  number of developers is beyond the scope of this paper.  However, developers
  can see the impact of a different weighting themselves.  Point to spreadsheet.
  Say what text is added to the paper.  Put it in blue?  \smallskip}

  \item In addition to github, could any other reference have been used, such as
  the number of citations in papers or relevant websites in the field? \medskip

  \emph{Citations and websites are a good idea, but don't have this data.
  Papers unreliable because people don't tend to cite scientific softwarwe.
  cite Katz to support this.  Changes to paper.\smallskip}

  \item Where are the quality ratings for each software? \medskip

  \emph{Double blind review meant didn't show citations, but citations have the
  full quality ratings.  The overall quality rating is given in Figure 2. The
  actual numbers are not as relevant as the ranking. \smallskip}

  \item Caption tables above them, not below them. \medskip

  \emph{This has been fixed in the revised manuscript.  \smallskip}

  \item What are the possible reasons for the differences between the ranking
  found and that of GitHub, in addition to the impossibility of installing some
  of them? \medskip

  \emph{Add content from the arxiv version.  ``Further reason for discrepancies
  between our ranking and the community's ranking is that we weighted all
  qualities equally. This is not likely how users implicitly rank the different
  qualities. As a result, some projects with high community popularity may have
  scored lower with our method because of a relatively higher (compared to the
  scientific community's implicit ranking) weighting of the poor scores for some
  qualities. A further explanation for discrepancies between our measures and
  the star measures may also be due to inaccuracy with using stars to
  approximate popularity.  Stars are not an ideal measure because stars
  represent the community's feeling in the past more than they measure current
  preferences \citep{Szulik2017}.  The issue with stars is that they tend only
  to be added, not removed.  A final reason for inconsistencies between our
  ranking and the community's ranking is that, as for consumer products, more
  factors influence popularity than just quality.'')\smallskip}

  \item Some software could not be installed. What problems were found? Why not
  remove it from the list? \medskip

  \emph{Say the problems from installability section.  Didn't remove because
  only affected a few qualities.  Favoured including with a warning.
  \smallskip}

  \end{enumerate}

  \textbf{Reviewer \#2}

  \begin{enumerate}

  \item The introduction section should contain the scope, significance of the
  research by summarizing current understanding and background information,
  stating the purpose of the work, and highlighting the potential outcomes.
  Also, there's no need to divide various subsection in the introduction.
  \medskip

  \emph{Response.\smallskip}

  \item The title of the table should be placed above the table. \medskip

  \emph{This has been fixed in the revised manuscript.  \smallskip}

  \item It is not recommended to include tables in the conclusion. \medskip

  \emph{Move it.  \smallskip}

  \end{enumerate}

  \closing{Best regards,~\newline} \vspace{-29mm}
  \includegraphics[scale=0.75]{Signature.pdf}

\end {letter}

\end{document}

### 2025-01-24 StateOfPractice/Papers/SOP_MI_CritReviewsBioMedEng/OrigSubmissionSOP_MI_OSR.tex
Preparing resubmission of SOP for MI for Crit Reviews in Bio Med Eng

We show the recommended artifacts in Table~\ref{Tbl_Guidelines}, one per row.
For a given row, a checkmark in one of the columns means that the corresponding
guideline recommends this artifact.  The last column shows whether the artifact
appears in the measured set of MI software, either not at all (blank), commonly
(C) (20 to 29 ($>$67\%) packages), uncommonly (U) (10 to 19 (33-67\%) packages)
or rarely (R) (1 to 9 ($<$33\%) packages).  We did our best to interpret the
meaning of each artifact consistently between guidelines and specific MI
software, but the terminology and the contents of artifacts are not
standardized.  The challenge even exists for the ubiquitous README file.  The
content of README files shows significant variation between
projects~\cite{PranaEtAl2018}.  Some content is reasonably consistent, with 97\%
of README files contain at least one section describing the `What' of the
repository and 89\% offering some `How' content, other categories are more
variable.  For instance, information on `Contribution', `Why', and `Who', appear
in 28\%, 26\% and 53\% of the analyzed files, respectively~\cite{PranaEtAl2018}.

\begin{table*}[ht!]
    \begin{center}
    \begin{tabular}{ p{3.8 cm} p{6.4 cm} p{4.5 cm}}
    \toprule
    Common & Uncommon & Rare \\
    \midrule
    README (29) & Build scripts (18) & Getting Started (9)\\
    Version control (29) & Tutorials (18) & Developer's manual (8)\\
    License (28) & Installation guide (16) & Contributing (8)\\
    Issue tracker (28) & Test cases (15) & API documentation (7)\\
    User manual (22) & Authors (14) & Dependency list (7)\\
    Release info. (22) & Frequently Asked Questions (14) & Troubleshooting guide (6)\\
     & Acknowledgements (12) & Product roadmap (5)\\
     & Changelog (12) & Design documentation (5)\\
     & Citation (11) & Code style guide (3)\\
     & & Code of conduct (1)\\
     & & Requirements (1)\\
    \bottomrule
    \end{tabular}
    \caption{Artifacts Present in MI Packages, Classified by Frequency
    (number of occurrences)}
    \label{artifactspresent}
    \end{center}
\end{table*}

Popularity in Table~\ref{Tbl_Guidelines} does not imply that these oft
recommended artifacts are the most important. Guidelines are often brief, to
encourage adoption, and thus even guidelines that mention the need for
installation instructions rarely mention uninstallation instructions.  Two items
in Table~\ref{artifactspresent}, which presents our measurements for the MI
software, do not appear in any guidelines: \emph{Troubleshooting guide} and
\emph{Developer's manual}.  However the information within these documents
overlaps with the recommended artifacts.  Troubleshooting information often can
be found in a User Manual, while the information in a ``Developer's Manual'' is
often scattered amongst many other documents.

@@

Neglecting requirements documentation is unfortunately common for research
software, and MI software is no exception to this trend.  Although such
documentation is recommended by some \cite{TobiasEtAl2018, HerouxEtAl2008,
SmithAndKoothoor2016}, in practice this is rare~\cite{HeatonAndCarver2015}.
Sanders and Kelly \cite{SandersAndKelly2008} interviewed 16 scientists from 10
disciplines and found that none of the scientists created requirements
specifications, unless regulations in their field mandated such a document.
Requirements are the least commonly produced type of documentation for research
software in general \cite{Nguyen-HoanEtAl2010}.

This is unfortunate as when scientific developers are surveyed on their pain
points, Wiese et al.~\cite{WieseEtAl2019} found that software requirements and
management is the software engineering discipline that most hurts them,
accounting for 23\% of the technical problems reported by study participants.
Further adding to the misfortune, there is a widespread perception that up-front
requirements are impossible for research software \cite{CarverEtAl2007,
SegalAndMorris2008}. Fortunately an iterative approach to requirements is
feasible~\cite{Smith2016}, and research-software specific templates
exist~\cite{SmithEtAl2007}.

@@

Other communities use checklists to help with best practices. Examples include
checklists for merging branches~\cite{Brown2015}, for saving and sharing changes
to the project~\cite{WilsonEtAl2016}, for new and departing team
members~\cite{HerouxAndBernholdt2018}, for processes related to commits and
releases~\cite{HerouxEtAl2008} and for overall software
quality~\cite{ThielEtAl2020, SSI2022}.

MI projects fall somewhat short of recommended best practices, but are not alone
amongst research software projects. This gap has been documented
before~\cite{Storer2017,JohansonAndHasselbring2018}, and is known to cause
sustainability and reliability problems \cite{FaulkEtAl2009}, and to waste
development effort~\cite{deSouzaEtAl2019}.

\section{Comparison of Tool Usage Between MI and Other Research Software}
\label{Sec_CompareTools}

We use data from interviews, as described in Section~\ref{sec_interview_methods}
to answer \rqref{RQ_CompareToolsProjMngmnt} (tool usage),
\rqref{RQ_CompareMethodologies} (methodology) and \rqref{RQ_PainPoints} (pain
points). Requests were sent to all 29 projects.  Nine developers from eight of
the projects agreed to participate: \textit{3D Slicer}, \textit{INVESALIUS 3},
\textit{dwv}, \textit{BioImage Suite Web}, \textit{ITK-SNAP},
\textit{MRIcroGL}, \textit{Weasis}, and \textit{OHIF}. We spent about 90
minutes for each interview. The full interview answers can be
found in [citation redacted for blind review].
%\cite{Dong2021}.

@@

\paragraph{Version control} Most teams interviewed (eight of nine) use GitHub
for version control, which makes it unsurprising that they also use pull
requests for managing community contributions. The experience of using GitHub
was reported as generally positive; some teams had actively migrated to GitHub
from other version control systems.  This represents a considerable uptake since
its rare use in 2006~\cite{Wilson2006}, which matches the uptake in other
communities: 10 years ago Nguyen et al~\cite{Nguyen-HoanEtAl2010} estimated that
only 50\% of research software projects used version control, and by 2018 this
had jumped to over 80\%~\cite{AlNoamanyAndBorghi2018} with many
communities~\cite{Smith2018} reaching 100\% already in 2018. In 2022, version
control use sits at over 95\%~\cite{CarverEtAl2022}.

@@

Tool use for MI software is similar to tool use for ocean modelling
software~\cite{JungEtAl2022}.  Both use tools for editing, compiling, code
management, testing, building, and project management, but differ in the
latter's use of Kanban boards for project management.

@@

This matches observations for research software in general.
Scientific developers naturally use an agile philosophy \cite{AckroydEtAl2008,
CarverEtAl2007, EasterbrookAndJohns2009, Segal2005, HeatonAndCarver2015}, or an
amethododical process \cite{Kelly2013}, or a knowledge acquisition driven
process~\cite{Kelly2015}.  A waterfall-like process can work for research
software~\cite{Smith2016}, especially if the developers work iteratively and
incrementally, but externally document their work as if they followed a
rational design process~\cite{parnas1986rational}.

@@

Pinto et al.~\cite{PintoEtAl2018} list some pain points that did not come up in
our conversations: interruptions while coding, scope bloat, lack of user
feedback, hard to collaborate on software projects, and aloneness.  Wiese et
al.~\cite{WieseEtAl2019} mentions two more: reproducibility, and software scope
determination.  [citation redacted for blind review] %\cite{SmithEtAl2024} 
adds lack of software experience, technical debt, and documentation.
We cannot conclude that these are not also issues with MI software due to our
small response size.

@@

We provide specific recommendations for future consideration -- these are not
criticisms of past and/or current development practices. We expand on some of
the ideas from our interviews, and bring in further ideas that could have a
noticeable impact. The ``new'' ideas are not novel to this paper but rather from
the software engineering and software carpentry domains, but these ideas have
not yet seen wide adoption for MI software.  We list the ideas roughly in the
order of increasing implementation effort.

@@

\begin{itemize}
    \item Elimination of headaches associated with a separate integration phase
    \cite{Fowler2006}, \cite[p.\ 20]{HumbleAndFarley2010}.
    \item Detection and removal of bugs \cite{Fowler2006} via
    automated testing.  
    \item Everyone is always working on a stable base that always passes all tests.
    \item If the CI system uses generators and linters, it will also have
    current documentation and clean code.
\end{itemize}

@@

These tools have many benefits: finding potential bugs, finding memory
leaks, improving performance, standardizing code with respect to formatting,
removing silly errors before code reviews, and catching potential security
issues~\cite{SourceLevel2022_Lint}. Most popular programming languages possess
such tools. This helps with \ppref{P_LackDevTime}, \qref{Q_Maintainability},
and correctness (\ppref{P_Correctness}).

The pervasive use of such tools to enforce documentation standards
partially explains the relatively higher quality of statistical tools that are
part of the Comprehensive R Archive Network (CRAN) [citation redacted for blind
review].  %\cite{SmithEtAl2018_StatSoft}.

@@

Rigorous inspection finds 60--65\% of latent defects on
average, and often tops 85\% in defect removal efficiency \cite{Jones2008}. The
success rate of code inspection is generally higher than most forms of testing,
which average between 30--35\% for defect removal efficiency
\cite{EbertAndJones2009, Jones2008}. For research software, Kelly and Shepard
\cite{KellyAndShepard2000} show a task based inspection approach can be
effective.

Due to improving code quality and increasing knowledge transfer, peer review
addresses the same pain points and qualities as linters (\ppref{P_LackDevTime},
\qref{Q_Maintainability} and \ppref{P_Correctness}). The benefits can be
increased by extending reviews to all artifacts, including documentation, build
scripts, test cases and the development process itself.

@@

Although the advice to modularize research software to handle
complexity is common \cite{WilsonEtAl2014, StewartEtAl2017, Storer2017},
specific guidelines on how to divide the software into modules is less
prevalent.  Not every decomposition is a good design for supporting change, as
shown by Parnas~\cite{Parnas1972a}.  For instance, a design with low cohesion and high
coupling \cite[p.\ 48]{GhezziEtAl2003} will make change difficult. Especially in
research software, where change is inevitable, designers need to produce a
modularization that supports change.  Ocean modelling software is currently
feeling the pain of not emphasizing modularization in legacy code
\cite{JungEtAl2022}.

@@

We follow Ampatzoglou et al.'s \cite{AmpatzoglouEtAl2019} analysis of
threats to validity in software engineering secondary studies.

@@

A study is reliable if repeating it by another researcher, using the same
methodology, would lead to the same results~\cite{RunesonAndHost2009}.
We identify the following threats:
\begin{enumerate*}
\item One individual does the measures for all packages. A different
evaluator might find different results, due to differences in abilities,
experience, and biases.
\item The measurements for the full set of packages took several months (of
elapsed time).  Over this time the software repositories may have changed and
the reviewer's judgement may have drifted.
\end{enumerate*}

The measurement process used has previously been shown to be reasonably
reproducible.
%% DOUBLE BLIND
%Smith et al.~\cite{SmithEtAl2016} 
[citation redacted] reports grading five software products by two reviewers.
Their rankings were almost identical. As long as each grader uses consistent
definitions, the relative comparisons in the AHP results will be consistent
between graders.

@@

Construct validity is when the adopted metrics represent what they are intended
to measure~\cite{RunesonAndHost2009}. We have identified the following
potential issues:
\begin{itemize}
\item We make indirect measurement of software qualities since meaningful direct
measures for qualities like maintainability, reusability and verifiability, are
unavailable.  We follow the usual assumption that developers achieve higher
quality by following procedures and adhering to standards \cite[p.\
112]{VanVliet2000}.
\item We could not do all measurements for some software as we could not
install or build \textit{dwv}, \textit{GATE}, and \textit{DICOM Viewer}. We
used a deployed on-line version for \textit{dwv}, a virtual machine version for
\textit{GATE}, but no alternative for \textit{DICOM Viewer}.
\item Robustness measurements involve only two pieces of data, leading to
limited variation in the robustness scores. Our measurement-time budget limited
what we could achieve here.
\item Our maintainability proxies (higher ratio of comments to source,
high percentage of closed issues) have not been validated.
\item While smaller modules tend to be easier to reuse, small modules are
not necessarily good modules, nor understandable modules.
\item The understandability measure relies on 10 random source code files, but
the 10 files will not necessarily be representative. 
\item Our overall AHP ranking makes the unrealistic assumption of equal
weighting.
\item We approximated popularity by stars and watches.
\item Table~\ref{Tbl_Guidelines} required judgement as
not all guidelines use the same names for artifacts that contain
essentially the same information.
\end{itemize}

@@

Internal validity means that discovered causal relations are trustworthy and
cannot be explained by other factors~\cite{RunesonAndHost2009}. We identify
the following:

\begin{itemize}
\item Our search (Section~\ref{SecWhatProjects}) could have missed a relevant
package.
\item Our methodology assumes that development activities
will leave a trace in the repositories, but this is not necessarily true. For
instance, we saw little evidence of requirements
(Section~\ref{Sec_CompareArtifacts}), but teams might keep this kind of
information outside their repos.
\item We interviewed a relatively small sample of 8 teams. Their pain points
(Section~\ref{painpoints}) may not be representative of the rest of their
community.
\end{itemize}

@@

If the results of a study can be generalized to other situations, then the
study is externally valid \cite{RunesonAndHost2009}.  In other words, we cannot
generalize our results if the development of MI software is fundamentally
different from other research software.  Although there are differences, like
the importance of data privacy for MI data, we found the approach to developing
[other software (citation redacted)]
%% DOUBLE BLIND
%LBM software \cite{SmithEtAl2024} 
and MI software to be similar.  Except for
the domain specific aspects, the trends observed in the current study are
similar to that for other research software.

@@

Our grading scores indicate that \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji} and \textit{OHIF Viewer} are the overall top four.  However, the
separation between the top performers and the others is not extreme.  Almost all
packages do well on at least a few qualities, as shown in
Table~\ref{topperformerstable}, which summarizes the packages ranked first and
second for each quality. Almost 70\% (20 of 29) of the software packages appear
in the top two for at least two qualities.  The only packages that do not appear
in Table~\ref{topperformerstable}, or only appear once, are \textit{Papaya},
\textit{MatrixUser}, \textit{MRIcroGL}, \textit{XMedCon}, \textit{dicompyler},
\textit{DicomBrowser}, \textit{AMIDE}, \textit{3DimViewer}, and
\textit{Drishti}.

>>>>

We show the recommended artifacts in Table~\ref{Tbl_Guidelines}, one per row.
The columns correspond to guidelines. A checkmark in a column means that the
guideline suggests the artifact in this row.  The final column shows the
frequency of the artifact in the measured set of MI software.  The frequency
ranges from not at all (blank), rarely (R) ($<$33\%), uncommonly (U) (33-67\%)
to commonly (C) ($>$67\%).  Due to a lack of standardization, the mapping
between specific MI software and guidelines was challenging, even for the
ubiquitous README file.  Prana et al.~\cite{PranaEtAl2018} show significant
variation in the content of README files between projects. Although 97\% of
README files contain at least one section describing the `What' and 89\% offer
content on `How', other content is more variable.  For example, information on
`Who', `Contribution', and `Why', appear in 53\%, 28\%, 26\% of the analyzed
files, respectively~\cite{PranaEtAl2018}.

\begin{table*}[ht!]
    \begin{center}
    \begin{tabular}{ p{3.8 cm} p{6.4 cm} p{4.5 cm}}
    \toprule
    Common & Uncommon & Rare \\
    \midrule
    README (29) & Build scripts (18) & Getting Started (9)\\
    Version control (29) & Tutorials (18) & Developer's manual (8)\\
    License (28) & Installation guide (16) & Contributing (8)\\
    Issue tracker (28) & Test cases (15) & API documentation (7)\\
    User manual (22) & Authors (14) & Dependency list (7)\\
    Release info. (22) & Frequently Asked Questions (14) & Troubleshooting guide (6)\\
     & Acknowledgements (12) & Product roadmap (5)\\
     & Changelog (12) & Design documentation (5)\\
     & Citation (11) & Code style guide (3)\\
     & & Code of conduct (1)\\
     & & Requirements (1)\\
    \bottomrule
    \end{tabular}
    \caption{Software Artifacts Found in MI Packages, Classified by Their Number
    of Occurrences}
    \label{artifactspresent}
    \end{center}
\end{table*}

Popularity in Table~\ref{Tbl_Guidelines} does not mean these 
artifacts are more important than other artifacts. Guidelines are often brief, to
encourage adoption, and thus even guidelines that mention the need for
installation instructions rarely mention uninstallation instructions.  Two items
in Table~\ref{artifactspresent}, which presents our measurements for the MI
software, do not appear in any guidelines: \emph{Troubleshooting guide} and
\emph{Developer's manual}.  However the information within these documents
overlaps with the recommended artifacts.  Troubleshooting information often can
be found in a User Manual, while the information in a ``Developer's Manual'' is
often scattered amongst many other documents.

@@

Neglecting requirements documentation is unfortunately common for research
software, and MI software is no exception to this trend.  Although such
documentation is recommended by some \cite{TobiasEtAl2018, HerouxEtAl2008,
SmithAndKoothoor2016}, in practice this is rare~\cite{HeatonAndCarver2015}.  In
interviews with 16 scientists from 10 disciplines, the scientists admitted that
they only wrote a requirements specification if the regulations in their field
mandated it~\cite{SandersAndKelly2008}.  For research software in general,
requirements are the least commonly written document~\cite{Nguyen-HoanEtAl2010}.

This is unfortunate since when scientific developers are
surveyed~\cite{WieseEtAl2019}, software requirements and management proves to be
their greatest pain points, accounting for 23\% of their technical problems.
Further adding to their misfortune, up-front requirements are often viewed to be
impossible for research software \cite{CarverEtAl2007, SegalAndMorris2008}.
Fortunately an iterative approach to requirements is feasible~\cite{Smith2016},
and research-software specific templates exist~\cite{SmithEtAl2007}.

@@

Other communities use checklists as part of best practices. For instance, there
are checklists for commits and releases~\cite{HerouxEtAl2008}, for team
turnover~\cite{HerouxAndBernholdt2018}, for branch merging~\cite{Brown2015}, for
sharing and saving project changes~\cite{WilsonEtAl2016}, and for software
quality~\cite{ThielEtAl2020, SSI2022}.

MI software does not follow all recommended best practices, but they are not
alone amongst research software projects. This gap has been documented
before~\cite{Storer2017,JohansonAndHasselbring2018}, and is known to cause
sustainability and reliability problems \cite{FaulkEtAl2009}, and to waste
development effort~\cite{deSouzaEtAl2019}.

\section{Comparison Between MI and Other Research Software for Tool Usage}
\label{Sec_CompareTools}

We interview data, as described in Section~\ref{sec_interview_methods}
to answer \rqref{RQ_CompareToolsProjMngmnt} (tool usage),
\rqref{RQ_CompareMethodologies} (methodology) and \rqref{RQ_PainPoints} (pain
points). Requests were sent to all 29 projects.  Nine developers from eight of
the projects agreed to participate: \textit{3D Slicer}, \textit{INVESALIUS 3},
\textit{dwv}, \textit{BioImage Suite Web}, \textit{ITK-SNAP},
\textit{MRIcroGL}, \textit{Weasis}, and \textit{OHIF}. We spent about 90
minutes for each interview. The full interview answers can be
found in [citation redacted for blind review].
%\cite{Dong2021}.

@@

\paragraph{Version control} Most teams interviewed (eight of nine) use GitHub
for version control, which makes it unsurprising that they also use pull
requests for managing community contributions. The experience of using GitHub
was reported as generally positive; some teams had actively migrated to GitHub
from other version control systems.  This represents a considerable uptake since
its rare use in 2006~\cite{Wilson2006}, which matches the uptake in other
communities: 10 years ago only 50\% of research software used version
control~\cite{Nguyen-HoanEtAl2010}, but by 2018 this jumped to over
80\%~\cite{AlNoamanyAndBorghi2018} with many communities~\cite{Smith2018}
reaching 100\% at this time. In 2022, version control use sits at over
95\%~\cite{CarverEtAl2022}.

@@

Tool use for MI software and ocean modelling software is
similar~\cite{JungEtAl2022}.  Both use tools for code and project management,
compiling and building, testing, and, editing. They differ in the more frequent
use of project management via Kanban boards for ocean modelling.

@@

This matches previous observations for other research software.  Scientific
software developers often use an agile philosophy \cite{AckroydEtAl2008,
CarverEtAl2007, EasterbrookAndJohns2009, Segal2005, HeatonAndCarver2015}, or an
amethododical process \cite{Kelly2013}, or a knowledge acquisition driven
process~\cite{Kelly2015}.  A waterfall-like process can work for research
software~\cite{Smith2016}, especially if the developers work iteratively and
incrementally, but externally document their work as if they followed a rational
design process~\cite{parnas1986rational}.

@@

Pinto et al.~\cite{PintoEtAl2018} list other pain points not mentioned by our
interviewees: scope bloat, aloneness, lack of user feedback, interruptions while
coding, and collaboration challenges. Two more are~\cite{WieseEtAl2019}:
reproducibility, and software scope determination.  [citation redacted for blind
review] %\cite{SmithEtAl2024} 
adds lack of software experience, documentation and technical debt.  Our small
sample size means that we cannot conclude that these are not also issues with MI
software.

@@

We provide specific recommendations for future consideration -- these are not
criticisms of past and/or current development practices. We expand on some of
the ideas from our interviews, and bring in further ideas that could have a
noticeable impact. The ``new'' ideas are not novel to this paper but rather from
the software engineering and software carpentry domains, but these ideas have
not yet seen wide adoption for MI software.  We sorted the list of ideas roughly
on increasing effort.

@@

\begin{itemize}
    \item Remove headaches caused by a separate integration phase
    \cite{Fowler2006}, \cite[p.\ 20]{HumbleAndFarley2010}.
    \item Detection and removal of bugs \cite{Fowler2006} via
    automated testing.  
    \item The base is constantly stable for everyone, always passing all tests.
    \item The code will be standardized and the documentation current, if the CI
    system uses generators and linters.
\end{itemize}

@@

These tools have many benefits: finding potential bugs, finding memory
leaks, improving performance, creating standard compliant code,
removing trivial errors before code reviews, and finding security
issues~\cite{SourceLevel2022_Lint}. Most popular programming languages possess
such tools. This helps with \ppref{P_LackDevTime}, \qref{Q_Maintainability},
and correctness (\ppref{P_Correctness}).

The pervasive use of such tools to enforce documentation standards partially
explains the relatively high quality of software in the Comprehensive R Archive
Network (CRAN) [citation redacted for blind review].
%\cite{SmithEtAl2018_StatSoft}.

@@

Latent defects can found on average at a rate of 60--65\% by rigorous
inspection, with this rate reaching as high as 85\% \cite{Jones2008}. The
success rate of code inspection is generally higher than via testing, since testing
averages around 30--35\% for the efficiency of defect removal \cite{EbertAndJones2009,
Jones2008}. For research software, Kelly and Shepard \cite{KellyAndShepard2000}
show a task based inspection approach can be effective.

Peer review addresses the same pain points and qualities as linters
(\ppref{P_LackDevTime}, \ppref{P_Correctness}, and \qref{Q_Maintainability}),
since they both focus on improving code quality and increasing knowledge
transfer.  The benefits can be increased by extending reviews to all artifacts,
including documentation, build scripts, test cases and the development process
itself.

@@

The advice to modularize scientific software to handle complexity is common
\cite{WilsonEtAl2014, StewartEtAl2017, Storer2017}; however, specific guidelines
on the criteria to use for modularization is rare.  Parnas~\cite{Parnas1972a}
shows that not every decomposition supports change.  For instance, a design with
high coupling and low cohesion will make change challenging~\cite[p.\
48]{GhezziEtAl2003}. Since change is frequent in research software, the best
modularizations will support change. The pain of not previously designing for
change is currently being felt by ocean modelling community~\cite{JungEtAl2022}.

@@

We follow Ampatzoglou et al.'s \cite{AmpatzoglouEtAl2019} analysis of
threats to validity in secondary studies of software engineering.

@@

Reliability means a study can be repeated by another researcher, using the same
methodology, and the same results will be achieved~\cite{RunesonAndHost2009}.
We identify the following threats:
\begin{enumerate*}
\item A single person measures all packages. A different evaluator might find
different results, due to differences in abilities, experience, and biases.
\item The measurements for the full set of packages took several months (of
elapsed time).  Over this time the software repositories may have changed and
the reviewer's judgement may have drifted.
\end{enumerate*}

The measurement process used has previously been shown to be reasonably
reproducible.
%% DOUBLE BLIND
%Smith et al.~\cite{SmithEtAl2016} 
[citation redacted] reports grading five software products by two reviewers.
Their rankings were almost identical. The relative comparisons in the AHP
results will be consistent between graders, as long as the evaluators are
consistent in their judgement.

@@

If the adopted metrics match their intended measure~\cite{RunesonAndHost2009},
then construct validity is achieved. We have identified the following potential
issues:
\begin{itemize}
\item We make indirect measurement of software qualities since meaningful direct
measures are unavailable for reusability, maintainability, verifiability, etc.
We assume that following procedures and adhering to standards leads to higher
quality \cite[p.\ 112]{VanVliet2000}.
\item We could not do all measurements for some software as we could not
install or build \textit{dwv}, \textit{GATE}, and \textit{DICOM Viewer}. We
used a deployed on-line version for \textit{dwv}, a virtual machine version for
\textit{GATE}, but no alternative for \textit{DICOM Viewer}.
\item Robustness measurements involve only two pieces of data, leading to
limited variation in the robustness scores. Our measurement-time budget limited
what we could achieve here.
\item Our maintainability proxies (higher ratio of comments to source,
high percentage of closed issues) have not been validated.
\item While smaller modules tend to be easier to reuse, small modules are
not necessarily good modules, nor understandable modules.
\item The understandability measure is based on a random sample of 10 code
files, which many not be representative of the entire code base. 
\item The overall AHP ranking makes the unrealistic assumption of equal
weighting between qualities.
\item We used stars and watches to approximate popularity.
\item Subjective judgement was required for Table~\ref{Tbl_Guidelines} since not
all guidelines use the same names for artifacts that contain essentially the
same information.
\end{itemize}

@@

The study is internally valid if the suggested causal relations are trustworthy
and not explainable by other factors~\cite{RunesonAndHost2009}. The identified threats to internal validity are:

\begin{itemize}
\item Relevant packages could have been missed in our search
(Section~\ref{SecWhatProjects}).
\item We assume that evidence of development activities
will be visible in the repositories, but this is not necessarily true. For
instance, we saw little evidence of requirements
(Section~\ref{Sec_CompareArtifacts}), but teams might keep this kind of
information outside their repos.
\item We interviewed a relatively small sample of 8 teams, which means their
identified pain points might not be representative (Section~\ref{painpoints}).
\end{itemize}

@@

If the results of a study can be generalized to other situations, then the study
is externally valid \cite{RunesonAndHost2009}.  In other words, we cannot
generalize our results if there is a fundamental difference between the
development of MI software and other research software.  Although there are
differences, like the importance of data privacy for MI data, we found the
approach to developing [other software (citation redacted)]
%% DOUBLE BLIND
%LBM software \cite{SmithEtAl2024} 
and MI software to be similar.  Except for
the domain specific aspects, the trends observed in the current study are
similar to that for other research software.

@@

Our grading scores indicate that \textit{3D Slicer}, \textit{ImageJ},
\textit{Fiji} and \textit{OHIF Viewer} are the overall top four.  However, the
separation between the top performers and the others is not extreme.  Almost all
packages do well on at least a few qualities, as shown in
Table~\ref{topperformerstable}, where we summarize the first and second ranked
packages for each quality. Almost 70\% (20 of 29) of the software 
is ranked in the top two for at least two qualities.  Packages that do not
appear in Table~\ref{topperformerstable}, or only appear once, are
\textit{Papaya}, \textit{MatrixUser}, \textit{MRIcroGL}, \textit{XMedCon},
\textit{dicompyler}, \textit{DicomBrowser}, \textit{AMIDE}, \textit{3DimViewer},
and \textit{Drishti}.
